---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# chchpd

The goal of the `chchpd` is to allow NZBRI researchers to access data from the 
Christchurch Longitudinal Parkinson's study. Rather than deal directly with 
data in spreadsheets or databases, functions like `import_participants()` and 
`import_motor_scores()` are provided, to abstract away dealing with the raw 
data source. Currently, these access data from shared online Google Sheets. 
Eventually, however, the functions will be re-written to pull the data from the 
nascent NZBRI REDCap Parkinson's database. This change under-the-hood should be 
invisible to the user, as the source of the data is abstracted behind the 
functions provided. Data is returned as a separate datafrme for each data set. 
Users are responsible for linking them together as required.

## Installation

This package is only of interest and utility internally at NZBRI and hence can't 
 be made available via CRAN (i.e. `install.packages('chchpd')` will never 
work. Instead, install `chchpd` from its development repository on Github as 
follows:

```{r gh-installation, eval = FALSE}
# install.packages('devtools')
devtools::install_github('nzbri/chchpd')
```
If `install_github()` is invoked subsequently, the package will be downloaded 
and installed only if the version on Github is newer than the one installed 
locally.

## Example

Records across the various tables must be joined using either `subject_id` as 
an index (e.g. for linking to the participant table, as it includes information,
such as sex, that is constant for a subject), or `session_id` (to join various 
measures gathered at approximately the same assessment session for a given 
participant, and which might change at different time points.

The 'glue' between participants and the data gathered about them is the 
'sessions' table. For each participant, this lists the various assessment 
sessions they have had. In the data, these sessions were often described by 
idiosyncratic labels, such as `999BIO_F2` (indicating the follow-up session two 
years after baseline recruitment session in the Progression study. But the same 
session might also have served as the baseline in the more selective PET study, 
and been labelled, say, `999BIO_PET0`. This often idiosyncratic labelling is 
cured by the subject session maping table, which would have a record for both 
the `999BIO_F2` and `999BIO_PET0` sessions, linking them to the same 
standardised session code, which has a form like `999BIO_2016-0-28`. When 
importing various data sources (like HADS or UPDRS), their idosyncratic session 
labels are replaced by this standardised form. Thus it is easy to join multiple 
tables togther systematically, as below. Often a key thing is to specify just a 
restricted set of sessions, by speciofying a partocular study to filter them by:

```{r example}
# basic example code
participants = import_participants()
sessions = import_sessions(from_study = 'PET')
np = import_neuropsyc()
updrs = import_motor_scores()

dat = right_join(participants, sessions, by = 'subject_id') %>% 
  left_join(np, by = 'session_id') %>% 
  left_join(updrs, by = 'session_id')
```
